---
title: "Summarizing Data"
author: "Sai Ananthula"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r setup, include FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(openintro)
data(nycflights)
```

### Exercise 1

Most flights are on time but there are certain flights that are late and this is evident in the histogram with the bin width of 15. 

```{r code-chunk-label}

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram()

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)

ggplot(data = nycflights, aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)

```

### Exercise 2

68 flight to SFO in February

```{r}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO",month == 2)

dim(sfo_feb_flights)

```



### Exercise 3

Most of the flight to SFO are on time and there actually enough flights that arrive early enough to SFO that the median is -2 (minutes) for February. 

```{r}

ggplot(data=sfo_feb_flights, aes(x=dep_delay)) +
  geom_histogram(binwidth = 5)

sfo_feb_flights %>%
  summarise(median_dd = median(dep_delay),
            n         =n())
 

```

### Excercise 4

AA has a very varied timeframe for arrival delays as their IQR is 32.8 which is double the next highest.
```{r}
sfo_feb_flights %>%
  group_by(carrier) %>%
  summarise(median_dd = median(dep_delay),
            iqr_dd = IQR(dep_delay),
            n_flights =n()) %>%
  print.data.frame(.,digits=2)
```
###  Exercise 5
The reason for factor is the categorize the data and make it easier to understand. The histogram below is simple and east to read. 
```{r}
nycflights %>%
  group_by(month) %>%
  ggplot(., aes(x=factor(month), y=dep_delay, fill=factor(month)) ) +
    geom_boxplot(outlier.colo="blue", outlier.size=2) +
    theme_dark()
```

### Exercise 6
The difference for choosing between mean and median when looking at what month to leave NYC comes down to consistency. A month with the lower mean is more likely to have a consistent departure time since the mean is the the average of all data in a set. This is in contrast to the median which is the middle value so this could be affected by a couple very late flights causing the median to shift to a worse delay departure or vice versa if flight leave early. It really comes down to whether want consistency and in most cases the months with a low mean delay departure time see like a good time to leave New York. 

### Exercise 7
I would fly out of LGA since its on time departure rate is the highest out of the three. 

```{r}

nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))

nycflights %>%
  group_by(origin) %>%
  summarise(ot_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ot_dep_rate)) %>%
  print.data.frame(., digits=3)

```

### Exercise 8
Airplane velocity seem to increase in a way similar to a logarithmic function but it level off around 1900/2000 miles but does increase for extremely long trips at the 5000+ mile range. There does seem to be some form of correlation between distance of a flight and the speed of the airplane howver data on flights between 2600 and 5000 miles is needed since the data set has very little so a solid conclusion cannot be drawn. 

```{r}
nycflights <- nycflights %>%
  mutate(avg_speed = distance / (air_time/60))


ggplot(nycflights, aes(x=distance,avg_speed)) + geom_point() + 
  labs(title = "Speed vs Distance")
```

### Exercise 9
I Would say about 50 minutes assuming dep_delay is in minutes because it seems like you still get to your destination in time. 
```{r}

nycflights$carrier  <- as.factor(nycflights$carrier)

nycflights %>% 
  filter(carrier == "AA" | carrier == "DL" | carrier == "UA") %>%
ggplot(aes(x = dep_delay, arr_delay ,color = carrier)) + geom_point()


```

